Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2016-12-17 Master Thesis/synopsis.tex
Add proof for 5.2 (iii)

\begin{lemma}\phantomsection\label{5.2}
\begin{enumerate}[(i)]
\item Let $f: A\to B$ be a homomorphism of abelian groups with finite kernel and cokernel. Then $A$ is free-by-finite iff $B$ is free-by-finite.
\item Let $0\to A'\stackrel{i}{\to}A\stackrel{p}{\to}A''\to 0$ be an exact sequence of abelian groups. If $A'$ and $A''$ are free-by-finite, then $A$ is free-by-finite. Conversely, if $A$ is free-by-finite, then $A'$ is free-by-finite. If additionally, $A'$ is finitely generated, then $A''$ is free-by-finite.
\item Let $A$ be free-by-finite. Then $\bigcap_{n\in\mathbb{N}}nA=0$.
\end{enumerate}
\end{lemma}

>>>>

\begin{lemma}\phantomsection\label{5.2}
\begin{enumerate}[(i)]
\item Let $f: A\to B$ be a homomorphism of abelian groups with finite kernel and cokernel. Then $A$ is free-by-finite iff $B$ is free-by-finite.
\item Let $0\longrightarrow A'\stackrel{i}{\longrightarrow}A\stackrel{p}{\longrightarrow}A''\longrightarrow 0$ be an exact sequence of abelian groups. If $A'$ and $A''$ are free-by-finite, then $A$ is free-by-finite. Conversely, if $A$ is free-by-finite, then $A'$ is free-by-finite. If additionally, $A'$ is finitely generated, then $A''$ is free-by-finite.
\item Let $A$ be free-by-finite. Then $\bigcap_{n\in\mathbb{N}}nA=0$.
\end{enumerate}
\end{lemma}

\begin{proof}
\begin{enumerate}[(i)]
\item Let $A$ be free-by-finite, i.e. $A = A' \oplus F$ for some finite group $A'$ and free group $F$. Since $\ker(f)\subset A$ is finite, it's actually contained in $A'$.

Now let $B$ be free-by-finite and $K = \ker(f)$. Since the injective map $A/K \to B$ induced by $f$ can be restricted to their torsion subgroups, we see that $(A/K)_\text{tor}$ is finite. Every $a\in A_\text{tor}$ can be written as $a = a' + k$ for some $a'\in A,\ k\in K$ such that $na'\in K$ for some $n\in\mathbb{N}$. Since both $(A/K)_\text{tor}$ and $K$ are finite, $A_\text{tor}$ is finite as well. TODO
\item TODO
\item It suffices to show this assertion for finite groups and for $A=\mathbb{Z}$. For $A$ finite we have $nA = 0$ for $n = \#A$. Since $\mathbb{Z}$ has no nilpotent elements, the intersection of all prime ideals of $\mathbb{Z}$ is already trivial. \qedhere
\end{enumerate}
\end{proof}
